Alternating Automata Modulo First Order Theories

### 2019-02-07 quantifiers/body.tex
ok

Consider, for instance the problem of proving that $\lang{A} \subseteq
\lang{B}$, where $A$ and $B$ are the timed automata in
Fig. \ref{fig:ta}. The automaton $A$ ($B$) has a single clock $x$
($y$) which is reset on the transition from $s_1$ to $s_2$ ($q_1$ to
$q_2$). Intuititvely, $A$ recognizes the sequences in which all events
are $a$ and the second event, if it occurs at all, it occurs after at
least $2$ time units. On the other hand, $B$ recognizes the sequences
in which there exists an event $a$ such that no event occurs after it
in exactly $1$ time unit. Formally, $\lang{A} = \set{a,t \mid t \in
  \real} \cup \set{a, t_1, \ldots a, t_n \mid t_1, \ldots, t_n \in
  \real,~ n \geq 2,~ t_1 \geq 2}$ and $\lang{B} = \set{\sigma_1, t_1
  \ldots \sigma_{i-1}, t_{i-1}, a, t_i, \ldots \sigma_n, t_n \mid
  \sigma_1, \ldots, \sigma_n \in \set{a,b},~ n \geq 1,~ t_1, \ldots,
  t_n \in \real,~ t_i-t_{i-1} \neq 1}$. It is proved\footnote{See
  \cite[Theorem 1]{AlurMadhusudan04} for a proof.}  that $\clang{B}$
cannot be recognized by a timed automaton, thus the inclusion cannot
be proved in the framework of timed regular languages, as $\lang{A}
\cap \clang{B} = \emptyset$.

@@

We have implemented a version of the IMPACT semi-algorithm
\cite{McMillan06} in a prototype tool, avaliable
online\footnote{\url{http://nts.imag.fr/images/e/ea/Cav19b.tar.gz}}. The
tool is written in Java and uses the Z3 SMT solver \cite{z3}, via the
JavaSMT interface \cite{javasmt}, for spuriousness and coverage
queries and also for interpolant generation. Table
\ref{tab:experiments} reports the size of the input automaton in
bytes, the numbers of Predicates, Variables and Transitions, the
result of emptiness check, the number of Expanded and Visited Nodes
during the unfolding and the Time in miliseconds.  The experiments
were carried out on a MacOS x64 - 1.3 GHz Intel Core i5 - 8 GB 1867
MHz LPDDR3 machine.

>>>>

Consider, for instance the problem of proving that $\lang{A} \subseteq
\lang{B}$, where $A$ and $B$ are the timed automata in
Fig. \ref{fig:ta}. The automaton $A$ ($B$) has a single clock $x$
($y$) which is reset on the transition from $s_1$ to $s_2$ ($q_1$ to
$q_2$). Intuititvely, $A$ recognizes the sequences in which all events
are $a$ and the second event, if it occurs at all, it occurs after at
least $2$ time units. On the other hand, $B$ recognizes the sequences
in which there exists an event $a$ such that no event occurs after it
in exactly $1$ time unit. Formally, $\lang{A} = \set{a, t_1, \ldots a,
  t_n \mid t_1, \ldots, t_n \in \real,~ n > 1 \Rightarrow t_1 \geq 2}$
and $\lang{B} = \set{\sigma_1, t_1 \ldots \sigma_{i-1}, t_{i-1}, a,
  t_i, \ldots \sigma_n, t_n \mid \sigma_1, \ldots, \sigma_n \in
  \set{a,b},~ n \geq 1,~ t_1, \ldots, t_n \in \real,~ 1 < i < n
  \Rightarrow t_i-t_{i-1} \neq 1}$. It is proved\footnote{See
  \cite[Theorem 1]{AlurMadhusudan04} for a proof.}  that $\clang{B}$
cannot be recognized by a timed automaton, thus the inclusion cannot
be proved in the framework of timed regular languages, as $\lang{A}
\cap \clang{B} = \emptyset$.

@@

We have implemented a version of the IMPACT semi-algorithm
\cite{McMillan06} in a prototype tool, avaliable
online\footnote{\url{https://gofile.io/?c=7zB3g7}}. The tool is
written in Java and uses the Z3 SMT solver \cite{z3}, via the JavaSMT
interface \cite{javasmt}, for spuriousness and coverage queries and
also for interpolant generation. Table \ref{tab:experiments} reports
the size of the input automaton in bytes, the numbers of Predicates,
Variables and Transitions, the result of emptiness check, the number
of Expanded and Visited Nodes during the unfolding and the Time in
miliseconds.  The experiments were carried out on a MacOS x64 - 1.3
GHz Intel Core i5 - 8 GB 1867 MHz LPDDR3 machine.
